Efficiency

### 2018-01-28 DraftExec.Rnw
With boxes in place

%  For caption width
\usepackage{graphicx,caption}
% For Centering figs
\usepackage[export]{adjustbox}
% For split figures
\usepackage{subcaption}
%  appendix
\usepackage[titletoc,page]{appendix}

\usepackage{draftwatermark}
\SetWatermarkText{DRAFT}
\SetWatermarkScale{5}
\SetWatermarkLightness{0.95}

@@

        \vspace{10cm}
        \textbf{\huge Efficiency}

@@

        \vspace{1.5cm}
        2018

@@

@@

The choice of the pumpjet propulsion system is quite probably the most significant new technology inclusion for Australia's future submarines.  Unlike other modern technolgoies such as Air Independent Propulsion and Lithium Ion Batteries, which are either in service or entering service in multiple other navies, no other navies have have adopted pumpjet technology for their conventional sumbarines, which is currently only used by nuclear submarines \parencite(coorey2016).

Despite acknowledging that the pumpjet may have an adverse impact on dived endurance \parencite{sbs2017}, Naval Group justifies the superiority of the pumpjet in the following terms: "In a confrontation between two otherwise identical submarines, the one with pump jet propulsion always has the tactical advantage."  No further explanation of this superiority has been provided in detail, but it is widely held that the pumpjet offered acoustic advantages which were highly regarded by those Australians who were responsible for selecting the design partner \parencite{stewart2016}.

The comparison between two otherwise identical submarines (though not necessarily limited to engaging just each other in combat), provides a useful and suitable framework for evaluating a single significant technology choice, where most other significantly novel or new technologies were apparently excluded from consideration \parencite{stanford2017}.  The conclusion of this report is that the following statement can be made with confidence:
\\
\\
\emph{In a comparison between two otherwise identical submarines, the one with the pump jet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurace, and a lower overall range, than the one with a conventional propeller.  This will confer a substantial tactical and strategic advantage on the conventionally propelled submarine in a very broad range of operational scenarios.}
\\
\\
Whilst it is broadly acknowledged that propellers have superior efficiency at lower speeds, but jets might have higher efficiency at higher speeds, the precise point of cross-over seemed unclear in public debate.  One informed commentator suggested that the crossover point was likely to be below or near the transit speed of the submarine, and that the pumpjet may offer enhanced range and endurance over the whole mission profile as a consequence \parencite{davies2017}.  This report finds no evidence of a pumpjet of the type used on submarine ever being recommended as superior for efficiency in general, and even closely related technologies for more commecial use not advocated below 25kt. During my research I encountered domain experts confident that a pumpjet could simply not feasibly be superior to a suitable propeller at speeds below 18kt.  Evidence of propellers remaining generally superior to waterjets and pumpjets in the speed ranges up to and even over 30kt is abundant.  Consequently, the possibility of some cross-over point leading to any kind of range or endurance advantage relevant for a conventional submarine can be safely ruled out.  The basis for claims made by public officials to the contrary, such as those by Rear Admiral Sammut \parencite{sammut2017}, must be called into question.

@@

@@

@@

Perhaps the most important relationship to understand to comprehend the operation of submarines is the relationship between speed and drag .  This is important because it sets out the fundamental framework as it applies to any submarine, regardless of propulsion type.  (It's also a pretty important for planes, cars, missiles, torpedoes, and basically everything else.)

@@

This has a profound impact on the operation and engineering of maritime vessels.  Reducing velocity has such a substantial reduction on hull drag that going slower is almost always a reasonable means of conserving fuel overall.  To the extent that time is non-critical, slower is always much much better.  It's for this reason that during the financial crisis, many cargo shipping companies adopted the practice of slow steaming \parencite{liang2014} in order to conserve fuel, despite this causing a range of possible new engineering issues which need to be accounted for in order to operate the engines at lower than normal power for sustained periods \parencite[8-10]{sanguri2012}. The fuel saving from operating even 30\% slower means that around half the total energy is required for a given journey.  Other incremental costs and inefficiences from operating 'off-design' are frequently outweight by such a dramatic reduction in overall power demand.

@@

@@

One may also note from the previous figure that for each line representing a given pitch, the efficiency falls precipitously towards zero around the point where $J$ reaches $P/D$. This is because this extreme of the curve represents a circumstance where the propeller is rotating, but not actually moving any water, since it is moving through the water at about that speed which corresponds to the exact helical path defined by the pitch being traced. This represents circumstances where a propeller encounters effectively zero load.  This might be the case when a eases it's power, and the momentum of the ship temporarily keeps the ship moving without the propeller exerting much or any net thrust, or where alternative propoulsion sources (sails, or other propellers) are keeping a speed higher than what the propeller actually has push backwards on water to maintain.  Since drag due to friction scales with velocity squared, very low speed do correspond to increasing advance ratios for a given propeller size, as would increases in propeller size for a given speed and pitch ratio. In practice however, under self-propuslion conditions for open propellers, slowing down towards does not elevate the advance ratio far enough past beyond peak efficiencies for overall efficiency to fall very far, as can be seen in Figure \ref{fig:EfficiencyBulten.png}

@@

@@

\begin{figure}

\begin{subfigure}[b]{0.9\textwidth}
\includegraphics[width=0.8\linewidth, center]{BMTEfficiency1.png}
\end{subfigure}
\begin{subfigure}[b]{0.9\textwidth}
\includegraphics[width=0.8\linewidth, center]{BMTEfficiency2.png}
\end{subfigure}

\caption{Studies on submerged waterjets for military surface craft currently show that the efficiency of jets decline sharply below 10kt \parencite{giles2010}}
\label{fig:BMTEfficiency.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=.8\textwidth}
\includegraphics[width=.8\textwidth]{JetHighFlow.png}
\caption{At high flow rates the water has sufficient velocity (indicated by arrow length) to overcome the adverse pressure gradients created by the duct shape, shown by the change from green to red. After the impeller has added further pressure to the water, the water accelerates out of the contracting nozzle through the favourable pressure gradient (red to green) to produce a jet, reaching equal pressure to outside at the exit}
\label{fig:JetHighFlow.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=.8\textwidth}
\includegraphics[width=.8\textwidth]{JetLowFlow.png}
\caption{At low flow rates the water has insufficient velocity to overcome the adverse pressure gradients created by the duct shape, and flow separation begins to occur on the duct intake, and between the blades beginning near the roots (stall).  Flow through the jet would be unsteady, with excessive energy expended creating vortices and turbulence that results in little thrust.}
\label{fig:JetLowFlow.png}
\end{figure}

@@

@@

\begin{figure}[h] \captionsetup{width=0.5\textwidth}
\includegraphics[width=\textwidth]{AcceleratingIdeal.png}
\caption{Calculations of 'ideal' efficiencies of accelerating ducts (neglecting drag) in \cite{oosterveld1970} show that a duct can only improve the efficiency of a propeller if it is loaded above a certain degree, and that longer ducts produce better improvements in 'ideal' efficiencies}
\label{fig:AcceleratingIdeal.png}
\end{figure}

@@

\caption{(a) A an intake may experience risks of flow separation even at low speeds (Inlet Velocity Ratio) on some surfaces due to certain flow deflections being more severe. (b) A an intake may experience risk of flow separation on different surfaces at high speeds (Inlet Velocity Ratio) where other flow deflections are more severe. Source: \cite{bulten2006numerical}}
\label{fig:LowIVR.png}
\end{figure}

@@

>>>>

%  For caption width
\usepackage{graphicx,caption}
% For Centering figs
\usepackage[export]{adjustbox}
% For split figures
\usepackage{subcaption}
%  appendix
\usepackage[titletoc,page]{appendix}
\setlength{\parindent}{0em}

\usepackage{tikz,lipsum,lmodern}
\usepackage[most]{tcolorbox}

\usepackage{draftwatermark}
\SetWatermarkText{DRAFT}
\SetWatermarkScale{2}
\SetWatermarkLightness{0.95}

@@

        \vspace{10cm}
        \textbf{\huge A comparison of pumpjets and propellers for non-nuclear submarine propulsion}

@@

        \vspace{1.5cm}
        January 2018

@@

\newpage
\setlength{\parskip}{1em}

@@

The choice of the pumpjet propulsion system is quite probably the most significant new technology inclusion for Australia's future submarines.  Unlike other modern technolgoies such as Air Independent Propulsion and Lithium Ion Batteries, which are either in service or entering service in multiple other navies, no other navies have have adopted pumpjet technology for their conventional sumbarines, which is currently only used by nuclear submarines \parencite(coorey2016).  Questions have been raised about the efficiency of pumpjets at low speeds \parencite{davies2017}. Propulsive efficiency at low speeds might be very important for conventional submarines since it will largely determine how long they can remain underwater before sufracing to recharge their batteries, but it would be of little or no concern to a nuclear submarine, which has effectively unlimited energy stores.

Despite acknowledging that the pumpjet may have an adverse impact on dived endurance \parencite{sbs2017}, Naval Group justifies the superiority of the pumpjet in the following terms: "In a contest between two otherwise identical submarines, the one with pump jet propulsion always has the tactical advantage."\parencite{guillou2017}  No further explanation of this superiority has been provided in detail.  This report does not attempt to engage with specific tactical encounters such as this, however, it is widely held that the pumpjet offered acoustic advantages which were highly regarded by those Australians who were responsible for selecting the design partner \parencite{stewart2016}.  Commentary from the designers of Shortfin Barracuda Block 1A seem to confirm that it is an acoustic advantage at elevated speeds which drive this tactical advantage when facing another submarine in combat \parencite{autret2016}.

The comparison between two "otherwise identical" submarines (though not necessarily limited to engaging just each other in combat), provides a useful and suitable framework for evaluating a potential trade-off in terms of other parameters, such as propulsive efficiency.  This is particularly the case given that most other significantly novel or new technologies which might dramatically impact propulsive performance were apparently not included for serious consideration \parencite[10]{stanford2017}.  Whilst many of the pages of this report are concerned with establishing the probable magnitude of the impact of efficiency at a range of speeds, the existence and direction of an efficiency difference between the two systems was quite firmly established at all the speeds most important for sumbarine operations. This allows for a surprisingly simple and strong qualitative conclusion to be reached in terms of the "otherwise identical" submarine comparison.

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= The Conclusive Qualitative Comparison]
In a comparison between two otherwise identical submarines, the one with the pumpjet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurace, and a lower overall range, than the one with a propeller.  This will confer a substantial tactical and strategic advantage on the conventionally propelled submarine in a very broad range of operational scenarios.
\end{tcolorbox}

Whilst it is broadly acknowledged that propellers have superior efficiency at lower speeds, but jets might have higher efficiency at higher speeds, the precise point of cross-over seemed unclear in public debate.  One informed commentator suggested that the crossover point was likely to be below or near the transit speed of the submarine, and that the pumpjet may offer enhanced range and endurance over the whole mission profile as a consequence \parencite{davies2017}.  This report finds no evidence of a pumpjet of the type used on submarine ever being recommended as superior for efficiency in general, and even closely related technologies for more commecial use not advocated below 25kt. During my research I encountered domain experts confident that a pumpjet could simply not feasibly be superior to a suitable propeller at speeds below 18kt.  Evidence of propellers remaining generally superior to waterjets and pumpjets in the speed ranges up to and even over 30kt is abundant.  Consequently, the possibility of some cross-over point leading to any kind of range or endurance advantage relevant for a conventional submarine can be safely ruled out.  The basis for claims made by public officials to the contrary, such as those by Rear Admiral Sammut \parencite{sammut2017}, must be called into question.

@@

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{HiFlowSmall.png}
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{LoFlowSmall.png}
\caption{}
\end{subfigure}

\caption{(a) At high speeds, the intake of a pumpjet slows the waterflow and raises its pressure, and it subsequently accelerates through the contracting nozzle to form an efficient jet (b) At low speeds, the pressure gradient created by the duct causes some of the waterflow to stop and reverse.  This 'flow separation' wastes energy spinning water in circles, and dramatically reduces the system's efficiency.}
\label{fig:HiLoSmall.png}
\end{figure}

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= Quantitative comparison under central assumptions]
\begin{center}
\begin{tabular}{ |c||c|c|c| }
 \hline
 \multicolumn{4}{|c|}{4kt Comparison Table} \\
 \hline
 Propulsion System & Submerged Range & Submerged Endurance &Indiscretion Ratio\\
 \hline
 Propeller &554nm & 138 hours & 2.9\%\\
 Pumpjet & 366nm & 91 hours & 4.4\%\\
 \hline
\end{tabular}

\vspace{6mm}
\begin{tabular}{ |c||c|c|c| }
 \hline
 \multicolumn{4}{|c|}{8kt Comparison Table} \\
 \hline
 Propulsion System & Submerged Range & Submerged Endurance & Indiscretion Ratio\\
 \hline
 Propeller &357nm & 45 hours & 9.6\%\\
 Pumpjet & 237nm & 30 hours & 15.1\%\\
 \hline
\end{tabular}
\vspace{6mm}

\begin{tabular}{ |p{3cm}|p{11cm}| }
 \hline
 \multicolumn{2}{|c|}{Assumption Table for "otherwise identical" submarines} \\
 \hline
 Batteries & 500 tonnes lead-acid, 0.14MJ/kg density, 100\% discharge\\
 Power Reference & 20kt reached by pumpjet submarine with 6.5MW power delivered\\
 Efficiency Curve & Central lines in Figure \ref{fig:Efficiency_assumptions} modeled on those in Figure \ref{fig:BMTEfficiency.png}\\
 Diesel Generators & 5MW delivered for effective re-charging\\
 Hotel Load & Constant 100kW\\
 AIP Technology & None\\
 \hline
\end{tabular}

\end{center}
\end{tcolorbox}

@@

\subsection{They physical relationship}
Perhaps the most important relationship to comprehend in the operation of submarines is the relationship between speed and drag .  This is important because it sets out the fundamental framework as it applies to any submarine, regardless of propulsion type.  (It's also a pretty important for planes, cars, missiles, torpedoes, and basically everything else.)

@@

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= Quick Summary: Sped and Drag]
Drag directly determines the energy required to cover a given distance. The physics relevant to submarines require that drag increases with speed squared, and the power required by a vessel to reach a given speed scales with speed cubed.  Consequently, the energy \textit{required} to move a vessel a given distance can vary massively over different speeds, often much more than the efficiency of any given system \textit{delivering that energy} will tend to vary.
\\
\\
This means that most vessels will be able to go much further by travelling extremely slowly, almost regardless of particulars of just how they are designed and what propulsion systems they use.  It also means that almost all the attention of propulsion system design focuses on 'design speed', generally pretty close to the top speed, since at very low speed so little thrust (hence fuel) is required that efficiency is of little concern.
\\
\\
For conventional submarines, however, the situation is very different.  They already adopt very low speeds (8-10kt) for transit to maximise the range they can get from their diesel.  Even most commercial shipping, including bulk carriers generally go much faster (16-24kt).  When a submarine dives, it relies on its batteries, which might store less than 1\% of the energy that's stored in its diesel tanks, and consequently has to slow down much more again (~4kt) to get meaningful range before returning to the surface to recharge. Whilst much higher speeds are possible even when dived, up to ~20kt, this has such a dramatic impact on the power demanded that it can't be sustained for useful periods of time, and are consequently only useful in an emergency, or very particular rapidly evolving tactical circumstances.
\\
\\
Since a conventional submarine already has to slow down to a crawl to get good dived range and endurance, the efficiency of the propulsion system now becomes of very great importance, as poor efficiency here will have a signifcant impact on how soon they have to return to the surface.  Nuclear submarines have effectively unlimited energy stores and don't require air to use it, and consequently don't share this concern.  They might also also spend much more of their dived time at much higher speeds as a result.
\end{tcolorbox}

\subsection{Practical implications for submarines}
This has a profound impact on the operation and engineering of maritime vessels.  Reducing velocity has such a substantial reduction on hull drag that going slower is almost always a reasonable means of conserving fuel overall.  To the extent that time is non-critical, slower is always much much better.  It's for this reason that during the financial crisis, many cargo shipping companies adopted the practice of slow steaming \parencite{liang2014} in order to conserve fuel, despite this causing a range of possible new engineering issues which need to be accounted for in order to operate the engines at lower than normal power for sustained periods \parencite[8-10]{sanguri2012}. The fuel saving from operating even 30\% slower means that around half the total energy is required for a given journey.  Other incremental costs and inefficiences from operating 'off-design' are frequently outweight by such a dramatic reduction in overall power demand.

@@

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= The other reason reactors are different]
Nuclear reactors have an enormous amount of energy stored in a small space, so for submarines they may not need refuelling for many years, if every.  However, they vary from other power plants in another crucial way, that can have downsides to their operations.
\\
\\
Most people take an internal combustion engine's ability to vary its power output for granted.  Simply open the throttle, and more power is produced, ease it right off and it drops again.  The engines that drive our cars and trucks are used to going from full power to almost no power every time we change gears, or stop in traffic.  Other than a tiny bit more long-term wear on the engine, there's little consequence to large power fluctuations.
\\
\\
Nuclear reactors don't enjoy the same luxury.  The fissioning of Uranium or Plutonium atoms leaves a range of other unstable nuclear isotopes as by-products, which in turn decay to other nuclear isotopes, some of which absorb neutrons in the reactor core, or have other impacts on the 'reactivity' of the reactor (how fast it goes). As a consequence, nuclear reactors can't freely oscilate thier power levels, as some of these by-products following sudden changes (particulalry dramatic reductions in power) can produce dangerous situations where the reactor can become unstable.  (It was following such a exercise in reducing power that the Chernobyl explosion occurred.)
\\
\\
The physics of these decay chains mean that nuclear reactors, including those in submarines generally avoid sudden changes in power.  Large reductions in power have to be taken in staged steps or a steady ramp-down over many hours, or more likely a couple of days.  Furthermore, if the reactor is ever completely shut-down, it generally can't be started safely again for some time, most likely a couple of days.  For obvious reasons, an operating sumbarine will strictly avoid anything that might risk requiring a shutdown.
\\
\\
The result is that nuclear submarines whenever operating at reduced speeds tend to have an abundance of spare power to dispose of.  The normal desire for energy 'efficiencies' are thrown into reverse, since it's undesirable to reduce the power of the reactor very far, and dangerous to reduce it fast. Whatever excess energy isn't sucked up by hotel load and propulsion has to be dispersed by additional systems, generally pumping water to dispose of the excess energy as heat. As such, in nuclear sumbarine design trading-off propulsive efficiency at low speeds is an easy, or even desirable choice.
\end{tcolorbox}

@@

One may also note from Figure \ref{fig:OpenWater.png} that for each line representing a given pitch, the efficiency falls precipitously towards zero around the point where $J$ reaches $P/D$. This is because this extreme of the curve represents a circumstance where the propeller is rotating, but not actually moving any water, since it is moving through the water at about that speed which corresponds to the exact helical path defined by the pitch being traced. This represents circumstances where a propeller encounters effectively zero load.  This might be the case when a eases it's power, and the momentum of the ship temporarily keeps the ship moving without the propeller exerting much or any net thrust, or where alternative propoulsion sources (sails, or other propellers) are keeping a speed higher than what the propeller actually has push backwards on water to maintain.  Since drag due to friction scales with velocity squared, very low speed do correspond to increasing advance ratios for a given propeller size, as would increases in propeller size for a given speed and pitch ratio. In practice however, under self-propuslion conditions for open propellers, slowing down towards does not elevate the advance ratio far enough past beyond peak efficiencies for overall efficiency to fall very far, as can be seen in Figure \ref{fig:EfficiencyBulten.png}

@@

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= The difference between a pumpjet and a propeller]
In it simplest form, a pumpjet can be thought of as a special type of ducted propeller, or propeller contained inside a shroud.  The impact of the shroud in a pumpjet has the special effect of \textit{decelerating} the water at the point that it reaches the propeller.  This has the dual effect of raising the pressure of the water at the point that the rotor has to do its work on it, which is how pumpjets are able to reduce or eliminate cavitation.
\\
\\
This type of duct shape contrasts with the system most commonly associated with ducted propellers on tug-boats and trawlers, known as accelerating ducts or 'Kort Nozzles'.  This type of duct performs essentially the opposite function to the decelerating duct, contracting in advance of the propeller and acclerating the flow at the stage that it meets propeller.  This effect accentuates the normal contraction of the water-column that occurs in open propellers, and can improve the efficiency of small propellers attempting to exert a large force at low speeds (such as tugs and trawlers) where an optimally sized propeller for such a role would be impractically large for the vessel.  In this circumstance the suction force on the inside of the duct helps generate a positive thrust that is greater than the negative drag on the exterior of the duct.
\end{tcolorbox}

@@

\begin{figure}[h]
\captionsetup{width = 0.9\textwidth}
\includegraphics[width=0.8\linewidth, center]{BMTEfficiency2.png}
\caption{Studies on submerged waterjets for military surface craft currently show that the efficiency of jets decline sharply below 10kt \parencite{giles2010}}
\label{fig:BMTEfficiency.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=.8\textwidth}
\includegraphics[width=.8\textwidth, center]{JetHighFlow.png}
\caption{At high flow rates the water has sufficient velocity (indicated by arrow length) to overcome the adverse pressure gradients created by the duct shape, shown by the change from green to red. After the impeller has added further pressure to the water, the water accelerates out of the contracting nozzle through the favourable pressure gradient (red to green) to produce a jet, reaching equal pressure to outside at the exit}
\label{fig:JetHighFlow.png}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=.8\textwidth}
\includegraphics[width=.8\textwidth, center]{JetLowFlow.png}
\caption{At low flow rates the water has insufficient velocity to overcome the adverse pressure gradients created by the duct shape, and flow separation begins to occur on the duct intake, and between the blades beginning near the roots (stall).  Flow through the jet would be unsteady, with excessive energy expended creating vortices and turbulence that results in little thrust.}
\label{fig:JetLowFlow.png}
\end{figure}

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= Why pumjet efficiency declines at low speed]
Pumpjets have lower efficiency at lower speeds because the shroud forces the stream of water to flow in a very different pattern to what it would naturally.  By expanding and decelerating the flow at the point where it meets the rotor (where it would contract and accelerate without the duct) the duct forces some of the flows kinetic energy to be converted back to pressure.
\\
\\
This has a terrible impact on efficiency when the incoming water doesn't have very much kinetic energy, and not a great deal more needs to be added to it, as happens when a vessel is moving slowly.  The effect of slowing down water that's already going slowly means that some of the water stops, and in fact turns around and creates a vortex, known as 'flow separation'.  The effective size of the water-colum being accelerated reduces and efficiency declines, as more of the system's energy is absorbed in spinning the water around in circles, rather than accelerating it backwards.
\end{tcolorbox}

@@

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= Can we just 'tune' the pumpjet to be more efficient at low speeds?]
A effects different effect and factors working in different directions make the process of re-designing a pumpjet quite complex, and bode poorly for a high-efficiency, low-speed pumpjet.  Generally increasing the mass-flow, by increasing duct size, might be assumed to be a first step. However this increases the size of the shroud, and efficiencies can decline further, since the duct inevitably produces some negative thrust.  Moderating the 'deceleration' by making the shape change more gradual could reduce some flow separation, but make the duct much longer, and increase overall friction drag from the surfaces.  Easy wins are not obvious.
\\
\\
Furthermore, with peak efficiencies for pumpjets generally currently well above 25kt, only quite radical design changes could be expected to have the sort of effect that might make a big enough difference (good efficiency below 10kt), which aren't really within the scope of 'tuning'.  The radical redesigns which might look capable of having a significant impact start to look very much like evolving right back to a propeller, which defeats the acoustic purposes of having a pumpjet in the first place.
\end{tcolorbox}

@@

\begin{figure}[h] \captionsetup{width=0.8\textwidth}
\includegraphics[width=\textwidth]{AcceleratingIdeal.png}
\caption{Calculations of 'ideal' efficiencies of accelerating ducts (neglecting drag) in \cite{oosterveld1970} show that a duct can only improve the efficiency of a propeller if it is loaded above a certain degree, and that longer ducts produce better improvements in 'ideal' efficiencies}
\label{fig:AcceleratingIdeal.png}
\end{figure}

@@

\caption{(a) An intake may experience risks of flow separation even at low speeds (Inlet Velocity Ratio) on some surfaces due to certain flow deflections being more severe. (b) A an intake may experience risk of flow separation on different surfaces at high speeds (Inlet Velocity Ratio) where other flow deflections are more severe. Source: \cite{bulten2006numerical}}
\label{fig:LowIVR.png}
\end{figure}

@@

\begin{tcolorbox}[width=0.8\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= What about hotel load and the system as a whole?]
The non-propulsive power demand of the submarine (hotel load) has an enormous impact on a conventional submarine's performance underwater. As extra power is drained as time ticks by, underwater endurance and range decline substantially with higher hotel loads.
\\
\\
Hotel load also has a significant impact on the \textit{difference} between the pumpjet and propeller affect the performance of the whole system. By driving the speed at which the propulsion becomes the dominant power demand higher, an extra-large hotel load reduces the overall difference between the two identical submarines, and pushes the peak proportional difference to higher speeds, where endurance and range is naturally lower due to drag increasing.
\\
\\
The inverse relationship leads to even more striking results, particularly if large energy stores (AIP or lithium-ion batteries) are available.  If the hotel load is very low (either because it was designed that way, or discretionary components are shut-down in a high-threat sitation to maximise endurance) then the penalty inflicted by the pumpjet becomes extremely large, quite probably close to a factor of two, amounting to around or over a week of endurance at 3kt, and many hundreds of nautical miles at slightly higher speeds.
\\
\\
The necessary consequence of this dynamic is that selecting a power-hungry combat system such as the AN/BYG-1 in conjunction with the pumpjet and lead-acid batteries for energy when dived will virtually ensure that many substantially smaller conventional sumbarines operating in the Asia Pacific in the coming decades will easily outclass the Shortfin Barracuda in terms of dived range and endurance.
\end{tcolorbox}
